Using Optimal Time Step Selection to Boost the Accuracy of Finite Difference Schemes for Time Dependent PDEs

### 2010-05-21 papers/OTSPDE/OTSPDE.tex
Added brief discussion of benefits of OTS selection for GPU computing.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Manuscript started: 2007/09/01
% First completed draft: 2008/05/12, Vitamin D, Inc.
% Submission draft: 2008/06/08, Vitamin D, Inc.
% Resubmission draft: 2008/11/16, Vitamin D, Inc.
% Resubmission draft: 2009/05/25, Serendipity Research
% Resubmission draft: 2010/03/25, Serendipity Research
% Resubmission draft: 2010/05/02, Serendipity Research
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{abstract}
In this article, we present a simple technique for boosting the order of 
accuracy of finite difference schemes for time dependent 
partial differential equations by optimally selecting the time step used
to advance the numerical solution. The power of the technique is its ability
to increase the accuracy of existing finite difference schemes with 
minimal modification to the original numerical algorithm.  Through
straightforward numerical analysis arguments, we explain the origin of the
boost in accuracy and estimate the computational cost of the resulting
numerical method.  We demonstrate the utility of optimal time step (OTS)
selection on several different types of finite difference schemes
for a wide array of classical linear and semilinear PDEs in one and more space
dimensions on both regular and irregular domains.  
\end{abstract}

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Manuscript started: 2007/09/01
% First completed draft: 2008/05/12, Vitamin D, Inc.
% Submission draft: 2008/06/08, Vitamin D, Inc.
% Resubmission draft: 2008/11/16, Vitamin D, Inc.
% Resubmission draft: 2009/05/25, Serendipity Research
% Resubmission draft: 2010/03/25, Serendipity Research
% Resubmission draft: 2010/05/21, Serendipity Research
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{abstract}
In this article, we present a simple technique for boosting the order of 
accuracy of finite difference schemes for time dependent 
partial differential equations by optimally selecting the time step used
to advance the numerical solution. The power of the technique is its ability
to increase the accuracy of existing finite difference schemes with 
minimal modification to the original numerical algorithm.  Through
straightforward numerical analysis arguments, we explain the origin of the
boost in accuracy and estimate the computational cost of the resulting
numerical method.  We demonstrate the utility of optimal time step (OTS)
selection on several different types of finite difference schemes
for a wide array of classical linear and semilinear PDEs in one and more space
dimensions on both regular and irregular domains.  OTS selection is 
particularly valuable in light of recent advances in GPGPU computing because
it significantly increases the usefulness of simple explicit time stepping
schemes, which are especially well-suited to GPU hardware.
\end{abstract}

@@

From a scientific computing perspective, recent advances in GPGPU computing
make optimal time step selection especially relevant.  Because explicit time
integration schemes are better suited to the GPU hardware architectures than
implicit time integration schemes, the ability of optimal time step selection
to dramatically boost the accuracy of simple explicit schemes makes it a
powerful tool for achieving high-performance on GPUs.
